A Mechanised Library for Generic (Co)Programming

### 2024-03-07 paper/paper.tex
draft of abstract

%TODO mandatory: add short abstract of the document
\begin{abstract}

\end{abstract}

>>>>

%TODO mandatory: add short abstract of the document
\begin{abstract}
  Generic programming and recursion schemes provide powerful abstractions for
reasoning about program equivalences and deriving optimisations. However,
formalizing such abstractions in Coq often requires compromises, such as
imposing performance penalties, requiring the assumption of additional axioms,
or introducing unsafe casts into extracted code (for example, \ocaml{Obj.magic}
in OCaml). This paper presents a fully axiom-free formalization of hylomorphisms
(a recursion scheme capturing divide-and-conquer algorithms) in Coq, along with
their algebraic laws. The formalization allows for the extraction of clean,
natural-looking OCaml code for a variety of recursive programs.
% while enabling
% formal reasoning about program equivalences using the laws of hylomorphisms.

  The key contribution of this paper is an axiom-free framework for generic
programming with recursion schemes in Coq, focusing on proving algebraic laws,
reasoning about program equivalences, and extracting safe, idiomatic code. With
this framework, we formalize a series of divide-and conquer, dynamic
programming, and mutual recursion programs as hylomorphisms. We demonstrate that
the extracted OCaml code for the programs formalized in our framework is
efficient, resembles code that a human programmer might write, and contains no
uses of \ocaml{Obj.magic}. We also present a machine-checked proof of the
well-known short-cut fusion optimsation.

  By avoiding undafe coersions in the extracted code, this work provides a solid
foundation for developing efficient, verified recursive programs with strong
guarantees about the safety and interoperability of the generated OCaml code.
\end{abstract}
